A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-08-06 survey.tex
iterations -> frames (to be more consistent with the branching factor part)

\begin{itemize}
\item They are {\em simultaneous move} games, where more than one player can issue actions at the same time. Additionally, these actions are {\em durative}, i.e. actions are not instantaneous, but take some amount of time to complete.
\item RTS games are ``real-time'', which actually means is that each player has a very small amount of time to decide the next move. Compared to Chess, where players may have several minutes to decide the next action, in StarCraft, the game executes at 24 iterations per second, which means that players can act as fast as every 42ms, before the game state changes.
\item Most RTS games are partially observable: players can only see the part of the map that has been explored. This is referred to as the {\em fog-of-war}.
\item Most RTS games are non-deterministic. Some actions have a chance of success.
\item And finally, the complexity of these games, both in terms of state space size and in terms of number of actions available at each decision cycle is very large. For example, the state space of Chess is typically estimated to be around $10^{50}$, heads up no-limit Texas holdem poker around $10^{80}$, and Go around $10^{170}$. In comparison, the state space of StarCraft in a typical map is estimated to be many orders of magnitude larger than any of those, as discussed in the next section.
\end{itemize}

@@

Another way  to measure the complexity of the game is  by looking at the branching factor, $b$, and the depth of the game, $d$, as proposed in~\cite{Gaby}, with a total game complexity of $b^d$. In Chess, $b \approx 35$ and $d \approx 80$. In more complex games, like Go, $b \approx 30$ to $300$, and $d \approx 150$ to $200$. In order to determine the branching factor in StarCraft when an AI plays it, we must have in mind, that the AI can issue actions simultaneously to as many units in the game as desired. Thus, considering that, in a typical game, a player controls between 50 to 200 units, the branching factor would be between $u^{50}$ and $u^{200}$, where $u$ is the average number of actions each unit can execute. 
{\color{blue}
Estimating the value of $u$ is not easy, since the number of actions a unit can execute is highly dependent on the context. Let us make the following assumptions: 1) at most 16 enemy units will be in range of a friendly unit (larger values are possible, but unlikely), 2) when an AI plays StarCraft, it only makes sense to consider movement in the 8 cardinal directions per unit (instead of assuming that the player can issue a ``move'' command to anywhere in the map at any point in time), 3) for ``build'' actions, we consider that SCVs (terran worker units) only build in their current location (otherwise, if they need to move, we consider that as first issuing a ``move'' action, and then a ``build''), and 4) let's consider only the {\em Terran} race. With those assumptions, units in StarCraft can execute between 1 (units like ``Supply Depots'', whose only action is to be ``idle'') to 43 actions (Terran ``Ghosts''), with typical values around 20 to 30. Now, if we have in mind that actions have cool-down times, and thus not all units can execute all of the actions at every frame, we can take a conservative estimation of about $10$ possible actions per unit per game frame. This results in a conservative estimate for the branching factor between $b \in [10^{50},10^{200}]$, only considering units (ignoring the actions buildings can execute). 
}
%Assuming a conservative value for $u$ of about $30$ (moving to nearby positions, attacking units in range, using special abilities, etc.), we obtain a branching factor of $b$ between  $30^{50}$ and $30^{200}$.
Now, to compute $d$, we simply consider the fact that typical games last for about 25 minutes, which results in $d \approx 36000$ (25 minutes $\times$ 60 seconds $\times$ 24 iterations per second).

>>>>

\begin{itemize}
\item They are {\em simultaneous move} games, where more than one player can issue actions at the same time. Additionally, these actions are {\em durative}, i.e. actions are not instantaneous, but take some amount of time to complete.
\item RTS games are ``real-time'', which actually means is that each player has a very small amount of time to decide the next move. Compared to Chess, where players may have several minutes to decide the next action, in StarCraft, the game executes at 24 frames per second, which means that players can act as fast as every 42ms, before the game state changes.
\item Most RTS games are partially observable: players can only see the part of the map that has been explored. This is referred to as the {\em fog-of-war}.
\item Most RTS games are non-deterministic. Some actions have a chance of success.
\item And finally, the complexity of these games, both in terms of state space size and in terms of number of actions available at each decision cycle is very large. For example, the state space of Chess is typically estimated to be around $10^{50}$, heads up no-limit Texas holdem poker around $10^{80}$, and Go around $10^{170}$. In comparison, the state space of StarCraft in a typical map is estimated to be many orders of magnitude larger than any of those, as discussed in the next section.
\end{itemize}

@@

Another way  to measure the complexity of the game is  by looking at the branching factor, $b$, and the depth of the game, $d$, as proposed in~\cite{Gaby}, with a total game complexity of $b^d$. In Chess, $b \approx 35$ and $d \approx 80$. In more complex games, like Go, $b \approx 30$ to $300$, and $d \approx 150$ to $200$. In order to determine the branching factor in StarCraft when an AI plays it, we must have in mind, that the AI can issue actions simultaneously to as many units in the game as desired. Thus, considering that, in a typical game, a player controls between 50 to 200 units, the branching factor would be between $u^{50}$ and $u^{200}$, where $u$ is the average number of actions each unit can execute. 
{\color{blue}
Estimating the value of $u$ is not easy, since the number of actions a unit can execute is highly dependent on the context. Let us make the following assumptions: 1) at most 16 enemy units will be in range of a friendly unit (larger values are possible, but unlikely), 2) when an AI plays StarCraft, it only makes sense to consider movement in the 8 cardinal directions per unit (instead of assuming that the player can issue a ``move'' command to anywhere in the map at any point in time), 3) for ``build'' actions, we consider that SCVs (terran worker units) only build in their current location (otherwise, if they need to move, we consider that as first issuing a ``move'' action, and then a ``build''), and 4) let's consider only the {\em Terran} race. With those assumptions, units in StarCraft can execute between 1 (units like ``Supply Depots'', whose only action is to be ``idle'') to 43 actions (Terran ``Ghosts''), with typical values around 20 to 30. Now, if we have in mind that actions have cool-down times, and thus not all units can execute all of the actions at every frame, we can take a conservative estimation of about $10$ possible actions per unit per game frame. This results in a conservative estimate for the branching factor between $b \in [10^{50},10^{200}]$, only considering units (ignoring the actions buildings can execute). 
}
%Assuming a conservative value for $u$ of about $30$ (moving to nearby positions, attacking units in range, using special abilities, etc.), we obtain a branching factor of $b$ between  $30^{50}$ and $30^{200}$.
Now, to compute $d$, we simply consider the fact that typical games last for about 25 minutes, which results in $d \approx 36000$ (25 minutes $\times$ 60 seconds $\times$ 24 frames per second).
